Agenda diversity and coverage bias: A quantitative approach to the agenda-setting theory

### 2018-08-21 bmc_article.tex
Formula de topico en funcion del tiempo corregida

\par We perform \emph{non-negative matrix factorization (NMF)} \cite{xu2003document, lee1999learning} on the document-term matrix (\emph{M}) in order to detect the main topics in the corpus. A topic is a group of similar articles which roughly talks about the same subject. 
\emph{NMF} is an algorithm which factorize the matrix \emph{M} into two matrices \emph{W} and \emph{H} (eq.\ref{eq:nmf}), with the property that all three matrices have no negative elements. This non-negativity makes the resulting matrices easier to inspect, and very suitable for topic detection \footnote{ Even though there are other techniques in topic detection, as for instance LDA (Latent Dirichlet Allocation), the NMF decomposition suites perfect for the kind of corpus we  have analyzed in this work (A detailed comparison of both, NMF and LDA methods could be found in Supplementary Material)}. 
\begin{equation}
M^{(d \times t)} \sim H^{(d \times k)} \cdot W^{(k \times t)}
\label{eq:nmf}
\end{equation}
Such as the resulting matrix \emph{H} has dimensions of number of documents per $k$ and matrix \emph{W} has dimensions of $k$ per number of terms, the number $k$ is therefore interpreted as the number of topics in the documents and it is a parameter that must be set before the factorization.
\par The matrix \emph{H} is the representation of the documents in the topic-space. Each row is normalized to unit $l_1$-norm so that their components can be viewed as a degree of membership of a given document in the set of topics. In particular, the index of the largest component tells us which is the most representative topic of the document.
On the other hand, \emph{W} gives the topics representation in the original term-space. The largest components of a row give the group of words which represent more accurately each topic (called keywords) and therefore an insight of what the topic is talking about.
Since the factorization of eq.\ref{eq:nmf} usually can not be made exactly, it is approximated by minimizing the reconstruction error, i.e. the distance between matrix \emph{M} and its approximated form $\tilde{M} = H \cdot W$. This was performed \emph{NMF} through the python module \emph{scikit-learn} \cite{scikit-learn}.
\par A time-dependent representation is obtained by defining the temporal profile of the topic $i$, $W_i(day)$, by eq.\ref{eq:topic_weight} where $l(j)$ is the number of words of the document $j$, $h_{ji}$ is the degree of membership of document $j$ on topic $i$ and $\delta$ is the Kronecker delta. Providing by the fact that each document's vector can have all non-zero components, it is allowed that a document contributes to more than one topics' weight.
In order to reduce noise, we apply a linear filter with a sliding window of size 3, and finally we normalize the temporal profiles in order to describe each newspaper agenda as a distribution over the topics' space which evolves over time. This last normalization prevent us against the differences in the number of articles that each newspaper publishes.

\begin{equation}
W_i(day) = \sum_j^d l(j) \cdot h_{ji} \cdot \delta_{d,day}
\label{eq:topic_weight}
\end{equation}

>>>>

\par We perform \emph{non-negative matrix factorization (NMF)} \cite{xu2003document, lee1999learning} on the document-term matrix (\emph{M}) in order to detect the main topics in the corpus. A topic is a group of similar articles which roughly talks about the same subject. 
\emph{NMF} is an algorithm which factorize the matrix \emph{M} into two matrices \emph{W} and \emph{H} (eq.\ref{eq:nmf}), with the property that all three matrices have no negative elements. This non-negativity makes the resulting matrices easier to inspect, and very suitable for topic detection \footnote{ Even though there are other techniques in topic detection, as for instance LDA (Latent Dirichlet Allocation), the NMF decomposition suites perfect for the kind of corpus we  have analyzed in this work (A detailed comparison of both, NMF and LDA methods could be found in Supplementary Material)}. 
\begin{equation}
M^{(d \times t)} \sim H^{(d \times k)} \cdot W^{(k \times t)}
\label{eq:nmf}
\end{equation}
Such as the resulting matrix \emph{H} has dimensions of number of documents per $k$ and matrix \emph{W} has dimensions of $k$ per number of terms, the number $k$ is therefore interpreted as the number of topics in the documents and it is a parameter that must be set before the factorization.
\par The matrix \emph{H} is the representation of the documents in the topic-space. Each row is normalized to unit $l_1$-norm so that their components can be viewed as a degree of membership of a given document in the set of topics. In particular, the index of the largest component tells us which is the most representative topic of the document.
On the other hand, \emph{W} gives the topics representation in the original term-space. The largest components of a row give the group of words which represent more accurately each topic (called keywords) and therefore an insight of what the topic is talking about.
Since the factorization of eq.\ref{eq:nmf} usually can not be made exactly, it is approximated by minimizing the reconstruction error, i.e. the distance between matrix \emph{M} and its approximated form $\tilde{M} = H \cdot W$. This was performed \emph{NMF} through the python module \emph{scikit-learn} \cite{scikit-learn}.
\par A time-dependent representation is obtained by defining the temporal profile of the topic $i$, $W_i(day)$, by eq.\ref{eq:topic_weight} where $l(j)$ is the number of words of the document $j$, $h_{ji}$ is the degree of membership of document $j$ on topic $i$, $d_j$ is the date of document $j$, and $\delta$ is the Kronecker delta. Providing by the fact that each document's vector can have all non-zero components, it is allowed that a document contributes to more than one topics' weight.
In order to reduce noise, we apply a linear filter with a sliding window of size 3, and finally we normalize the temporal profiles in order to describe each newspaper agenda as a distribution over the topics' space which evolves over time. This last normalization prevent us against the differences in the number of articles that each newspaper publishes.

\begin{equation}
W_i(day) = \sum_j l(j) \cdot h_{ji} \cdot \delta_{d_j,day}
\label{eq:topic_weight}
\end{equation}
